---
title: Scraping Archived Data with the Wayback Machine
author: Peter Baumgartner
date: '2019-08-01'
lastmod: '2021-05-20'
categories:
  - API
  - open-science
  - web-service
tags:
  - wayback machine
  - web archive
slug: scraping-archived-data-with-the-wayback-machine
summary: 'In a previous article, I wrote about the possibilities of the Wayback
  Machine for scientific writing. I argued that archiving web pages prevent link rots
  when cited web resources are not available anymore. With this blog entry, I am looking
  quasi into the reverse option -- How to find and retrieve archived web pages for research
  reasons?'
commentable: true
draft: no
side_toc: false
output:
  blogdown::html_page:
    toc: true
    toc_depth: 4
image:
  placement: 2    
  caption: ""
  alt_text: ""
  focal_point: "Center"
  preview_only: false
---

## Why is archiving web pages important?

In a [previous article](/2019/07/22/archiving-quoted-web-resources/), I wrote about the possibilities of the [Wayback Machine](https://archive.org/web/) for scientific writing. I argued that archiving web pages are essential for references as they prevent link rots when cited web resources are not available anymore. With this blog entry, I am looking quasi into the reverse option: How to find and retrieve archived web pages for research reasons?

Archives web pages as permanently stored data are indispensable for reproducibility issues. But they are also valuable research resources as they are data for historical and comparative research. I will demonstrate the research significance with the historical analysis of [static website generators](https://www.staticgen.com/).

{{% callout note %}} Update: 2021-05-20: The above URL <https://www.staticgen.com/> recently redirects to <https://jamstack.org/generators/>. {{% /callout %}}

This here is the first part and shows how to use the Wayback Machine for retrieving archived web pages. The second part displays the results of the analysis which would not be possible without web archiving.

The nitty-gritty of this article comes from the excellent work of [Bob Rudis](https://rud.is/), who wrote many well documented [Tools to Work with the Various Internet Archive Wayback Machine APIs](https://github.com/hrbrmstr/wayback).

## Three different APIs for the Wayback Machine

There are [three different APIs](https://archive.org/help/wayback_api.php) for the Wayback Machine:

1.  **Wayback Availability JSON API**: A simple API for Wayback to test if a given url is archived and currenlty accessible in the Wayback Machine.
2.  **Memento API**: The Internet Archive Wayback Machine is also fully compliant with the [Memento Protocol](http://mementoweb.org/) which provides additional interfaces for querying snapshots (eg 'Mementos').
3.  **Wayback CDX Server API:** It allows for complex querying, filtering and analysis of Wayback capture data. See for details: [Wayback CDX Server \@ GitHub](https://github.com/internetarchive/wayback/tree/master/wayback-cdx-server).

I will explain the first two. The CDX Server is for complex querying, filtering and analysis and I didn't use it for my example.

## Three preparatory exercises

### Does the Internet Archive have my research URL cached?

At first you have to install the [wayback package](https://github.com/hrbrmstr/wayback) and load it so we can use Bob's library of functions.

{{% callout warning %}} Update: 2021-05-20: In the meanwhile there exists [another package with the same name on CRAN](https://cran.r-project.org/web/packages/wayback/index.html) (Comprehensive R Archive Network). It is dedicated to the installation for legacy R versions and has nothing to do with the wayback package on GitHub for the Wayback Machine of the Internet Archive. {{% /callout %}}

```{r install-packages}
if (!require("wayback"))
        {remotes::install_github("hrbrmstr/wayback", build_vignettes = TRUE)
        library(wayback)}
if (!require("reactable"))
        {install.packages("reactable")
        library(reactable)}
```

{{% callout note %}} Update 2021-05-20: For a better presentation of the example results I have also loaded the [reactable package](https://cran.r-project.org/web/packages/reactable/index.html). {{% /callout %}}

The URL I am looking for is: <https://www.staticgen.com/>. I am going to use `archive_available(url, timestamp)`: The timestamp parameter is optional. If it is lacking then the query date is used. If the URL is stored then the function returns the chronologically nearest archived version. Return value is a tibble with one observation and 5 variables:

```{r url-cached}
staticgen_avail <- archive_available("https://www.staticgen.com/")
reactable(staticgen_avail)
```

### Retrieve site mementos from the Internet Archive

Mementos are prior versions of web pages cached by web crawlers and stored in web archives. The [Internet Archive](https://archive.org/)) is one of these web archives but there exist other software as well, including systems that support versioning such as wikis or revision control systems. The HTTP-based Memento framework is a [description for Time-Based Access to Resource States](https://mementoweb.org/guide/rfc/).

> The HTTP-based Memento framework bridges the present and past Web. It facilitates obtaining representations of prior states of a given resource by introducing datetime negotiation and TimeMaps. Datetime negotiation is a variation on content negotiation that leverages the given resource's URI and a user agent's preferred datetime. TimeMaps are lists that enumerate URIs of resources that encapsulate prior states of the given resource. The framework also facilitates recognizing a resource that encapsulates a frozen prior state of another resource.

There are several resources for a better understanding of the memento framework:

-   An introductory [guide to the memento solution](https://mementoweb.org/guide/quick-intro/)
-   An interactive [Time Travel search engine](https://timetravel.mementoweb.org/)
-   There are many [memento tools](https://mementoweb.org/tools/) already available. Most important for the general user are browser extensions for Google Chrome and Firefox.
-   There exists an [awesome-memento list](https://github.com/machawk1/awesome-memento) at Github with references to tons of resources.
-   And last but not least there is the formal description as [Request for Comments](https://mementoweb.org/guide/rfc/) (RFC 7089), from which I have taken the quote above.

The [Open Wayback software](https://web.archive.org/) used by the [Internet Archive](https://archive.org/) is fully compliant with RFC 7089, the specification of the Memento protocol.

With `get_mementos(url, timestamp = format(Sys.Date(), "%Y"))` we will receive a short list of relevant links to the archived content. The function returns the four link relation types as in the [Request for Comment for the Memento framework](https://mementoweb.org/guide/rfc/#Link-Header-Relation-Types) outlined.

1.  Link Relation Type "original"
2.  Link Relation Type "timemap"
3.  Link Relation Type "timegate"
4.  Link Relation Type "memento"

Besides these 4 main types of link relations the function provides also the first, previous and last available memento. Normally the last memento is identical with the memento link relation type. In addition to the two columns `link` and `rel` there is a third one `ts`, containing the time stamps (empty for the first three link relation types). The return value in total is a tibble with seven observations (rows) and three columns.

{{% callout note %}} Update 2021-05-20: Because of the redirection of the (old) URL the `get_mementos` function without `timestamp` parameter would not work. Even if the latest staticgen.com URL was archived only 16 days ago! {{% /callout %}}

```{r get-mementos}
staticgen_mntos <- get_mementos("https://www.staticgen.com/", timestamp = "2021")
reactable(staticgen_mntos)
```

### Get the point-in-time memento crawl list

Providing an URL in the search field of the Wayback Machine results in the interactive browser version to the [calender view](https://web.archive.org/web/*/https://www.staticgen.com). The dates with archived content are blue or green (redirected URL) circled. The bigger the circles the more snapshots were archived on these dates.

We get these dated crawl list with the second observation of the `get_mementos` function. The execution of the next code chunk can take several moments, depending how many pages of the URL are archived. Be aware that the Wayback server is strained by this query, so do not repeat it several times but store the result on your hard disk.

```{r get-timemap}
staticgen_tm <- get_timemap(staticgen_mntos$link[2])
reactable(staticgen_tm)
```

Included into the `r nrow(staticgen_tm)` captures of the interactive browser version there are four rows, relating to the four link relation types mentioned above. The last line is empty.

## Summary: Putting all together

We can put together all previous preliminay steps into a new function e.g., `get_crawl(url)`. This functions gets an URL and returns a list of all archived versions for this URL.

-   Check if for the URL exists an archived version. If not: stop execution.
-   If exists an archived version, then retrieve mementos for this url from the Internet Archive.
-   Get the point-in-time memento crawl list for this URL.
-   Clean up so that only memento links remain
-   Delete unnecessary rows `type` and `from`.
-   Convert row `datetime` from class 'character' to class 'datetime'.
-   Delete duplicate datetime records. (Sometimes there are more than one capture taken at the same day, refering to the URL and the port used.)
-   Filter rows with an algorithm, so that only those mementos remain which are suitable for the comparison analysis. For instance: Take the first memento for every year, or every month etc.

As these steps are either already illustrated or not specific to the Wayback Machine, I will stop here with my explications. I have provided an [article on RPubs](https://rpubs.com/pbaumgartner/wayback) with all the details. But keep in mind that I am not very experienced in R and there may be much simpler and more elegant solutions.

<div class="notices info" >
<p class="centered">
Read my [RPubs article](https://rpubs.com/pbaumgartner/wayback) with all the details of retrieving and web scrapping. Visit also my tutorial on [How-to use the Wayback Machine interactively?](/slide/wayback-machine-tutorial/).
</div>

<span class='Z3988' title='url_ver=Z39.88-2004&amp;ctx_ver=Z39.88-2004&amp;rfr_id=info%3Asid%2Fzotero.org%3A2&amp;rft_val_fmt=info%3Aofi%2Ffmt%3Akev%3Amtx%3Adc&amp;rft.type=blogPost&amp;rft.title=Scraping%20Archived%20Data%20with%20the%20Wayback%20Machine&amp;rft.source=Thought%20splinters&amp;rft.rights=CC%20BY-SA%204.0&amp;rft.description=In%20a%20previous%20article,%20I%20wrote%20about%20the%20possibilities%20of%20the%20Wayback%20Machine%20for%20scientific%20writing.%20I%20argued%20that%20archiving%20web%20pages%20prevent%20link%20rots%20when%20cited%20web%20resources%20are%20not%20available%20anymore.%20With%20this%20blog%20entry,%20I%20am%20looking%20quasi%20into%20the%20reverse%20option%20--%20How%20to%20find%20and%20retrieve%20archived%20web%20pages%20for%20research%20reasons?&amp;rft.identifier=https%3A%2F%2Fnotes.peter-baumgartner.net%2F2019%2F08%2F01%2Fscraping-archived-data-with-the-wayback-machine&amp;rft.aufirst=Peter&amp;rft.aulast=Baumgartner&amp;rft.au=Peter%20Baumgartner&amp;rft.date=2019-08-01&amp;rft.language=en'></span>
